---
title: "p8105_hw2_hg2596"
output: github_document
date: "2022-10-06"
---

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
library(readr)
```

# Problem 2 

# Import Mr.Trash Wheel 

```{r}
mtw = read_excel("./data/Trash_Wheel_Collection_Data.xlsx", sheet = "Mr_Trash_Wheel") 
```

# Clean data for Mr.Trash Wheel 

```{r}

mtw_2 = read_excel("./data/Trash_Wheel_Collection_Data.xlsx", sheet = "Mr_Trash_Wheel",range = cell_cols("A:N"), col_names = TRUE, col_types = NULL, trim_ws = TRUE,skip = 1, n_max = 535) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>%
  mutate(dumpster = as.character(dumpster)) %>%
  mutate(year = as.character(year)) %>%
  mutate(sports_balls = as.integer(sports_balls)) %>%
  mutate(marker = 1) %>%
  select(dumpster,marker,everything())
```

# Import Professor Trash Wheel 

```{r}
ptw = read_excel("./data/Trash_Wheel_Collection_Data.xlsx", sheet = "Professor_Trash_Wheel") 
```

# Clean data for Professor Trash Wheel 

```{r}
ptw_2 = read_excel("./data/Trash_Wheel_Collection_Data.xlsx", sheet = "Professor_Trash_Wheel", range = cell_cols("A:M"), col_names = TRUE, col_types = NULL, trim_ws = TRUE,skip = 1, n_max = 115) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(month)) %>%
  drop_na(dumpster)%>%
  mutate(dumpster = as.character(dumpster)) %>%
  mutate(year = as.character(year)) %>%
  mutate(marker = 2) %>%
  select(dumpster,marker,everything())
```

# Merge through joint 

```{r}
merge = bind_rows(mtw_2,ptw_2) 
```

# Find Professor Trash Wheel Weight of trash
```{r}
merge %>%
  filter(marker == 2) %>%
  pull(weight_tons) %>%
  sum(na.rm = TRUE) 
```

# Find sum of sports balls 
```{r}
merge %>%
  filter(year == 2020) %>%
  pull(sports_balls) %>%
  sum(na.rm = TRUE) 
```

This new dataset has `r nrow(merge)` observations and `r ncol(merge)` variables. Some key variables are **dumpster**, **month**, **date**, **year**, **weight(tons)**,**sports balls** and various other trash items. The **marker** variable is too help keep track of each dataset, "1" indicates it came from Mr.Trash Wheel dataset and "2" indicates it came from Professor Trash Wheel dataset. 

The total weight of trash collected in Professor Trash Wheel is `r merge %>% filter(marker == 2) %>% pull(weight_tons) %>% sum(na.rm = TRUE)` tons.` 

Also, the total number of sport balls `r merge %>% filter(year == 2020) %>% pull(sports_balls) %>% sum(na.rm = TRUE)`. 
