---
title: "p8105_hw2_hg2596"
output: github_document
date: "2022-10-06"
---

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
library(readr)
```

# Problem 1 

#Import data and clean dataset 

```{r}
trans_ent = 
  read_csv(
    "./data/NYC_data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

In the beginning of cleaning this dataset it's important to switch the **Route** variable into character for the last 8-11 since 1-7 are also character variables.

# Select and Disctict

```{r}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```

Converted the variable route from wide to long format, this helps us know things about specific routes.In the following step we use distinct function to help provide unique combinations.

# Find ADA complaint stations 

```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

In order to answer how many ADA complaint stations, we filter out for ADA when true and use distinct to tell us how many are complaint. 

# Proportions without vending machine 

```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

In order to figure out which proportion of the station entrances and exits without vending machines we filterd otu for vending machine when no, and pulled out the column for enter to find the mean to be ` r trans_ent %>%filter(vending == "NO") %>% pull(entry) %>% mean`. 

#Distinct Stations 

```{r}
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

In order to get distinct variables, we needed to use pivot longer. Then pull from our previous step to filter out the route for A and ada. 


# Problem 2 

# Import Mr.Trash Wheel 

```{r}
mtw = read_excel("./data/Trash_Wheel_Collection_Data.xlsx", sheet = "Mr_Trash_Wheel") 
```

# Clean data for Mr.Trash Wheel 

```{r}

mtw_2 = read_excel("./data/Trash_Wheel_Collection_Data.xlsx", sheet = "Mr_Trash_Wheel",range = cell_cols("A:N"), col_names = TRUE, col_types = NULL, trim_ws = TRUE,skip = 1, n_max = 535) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>%
  mutate(dumpster = as.character(dumpster)) %>%
  mutate(year = as.character(year)) %>%
  mutate(sports_balls = as.integer(sports_balls)) %>%
  mutate(marker = 1) %>%
  select(dumpster,marker,everything())
```

# Import Professor Trash Wheel 

```{r}
ptw = read_excel("./data/Trash_Wheel_Collection_Data.xlsx", sheet = "Professor_Trash_Wheel") 
```

# Clean data for Professor Trash Wheel 

```{r}
ptw_2 = read_excel("./data/Trash_Wheel_Collection_Data.xlsx", sheet = "Professor_Trash_Wheel", range = cell_cols("A:M"), col_names = TRUE, col_types = NULL, trim_ws = TRUE,skip = 1, n_max = 115) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(month)) %>%
  drop_na(dumpster)%>%
  mutate(dumpster = as.character(dumpster)) %>%
  mutate(year = as.character(year)) %>%
  mutate(marker = 2) %>%
  select(dumpster,marker,everything())
```

# Merge through joint 

```{r}
merge = bind_rows(mtw_2,ptw_2) 
```

# Find Professor Trash Wheel Weight of trash
```{r}
merge %>%
  filter(marker == 2) %>%
  pull(weight_tons) %>%
  sum(na.rm = TRUE) 
```

# Find sum of sports balls 
```{r}
merge %>%
  filter(year == 2020) %>%
  pull(sports_balls) %>%
  sum(na.rm = TRUE) 
```

This new dataset has `r nrow(merge)` observations and `r ncol(merge)` variables. Some key variables are **dumpster**, **month**, **date**, **year**, **weight(tons)**,**sports balls** and various other trash items. The **marker** variable is too help keep track of each dataset, "1" indicates it came from Mr.Trash Wheel dataset and "2" indicates it came from Professor Trash Wheel dataset. 

The total weight of trash collected in Professor Trash Wheel is `r merge %>% filter(marker == 2) %>% pull(weight_tons) %>% sum(na.rm = TRUE)` tons.Also, the total number of sport balls `r merge %>% filter(year == 2020) %>% pull(sports_balls) %>% sum(na.rm = TRUE)` from Mr.Trash in 2020.

# Problem 3 

# Upload Pols Month dataset
```{r}
pols_month = read_csv("./data_2/fivethirtyeight_datasets/pols_month.csv", show_col_types = FALSE)
```

# Clean Pols Month dataset and add Variables 

```{r}
pols_month_2 = read_csv("./data_2/fivethirtyeight_datasets/pols_month.csv", show_col_types = FALSE) %>% 
  janitor::clean_names() %>%
  separate(mon,c("A", "B", "C")) %>%
  select("B","C","A", everything()) %>%
  rename(month = "B", day = "C", year = "A") %>%
  mutate(month = month.abb[as.numeric(month)]) %>%
  mutate(president = case_when(prez_dem == 1 ~ "dem" , 
                   prez_gop > 0 ~ "gop",
                   TRUE ~ "")) %>%
  select(-prez_dem) %>%
  select(-prez_gop) %>%
  select(-day)%>%
  mutate(month = as.character(month)) %>%
  select(year,month,everything())
```

# Upload snp dataset 

```{r}
snp = read_csv("./data_2/fivethirtyeight_datasets/snp.csv", show_col_types = FALSE)
```

# Clean snp dataset and add variables 

```{r}
snp_2 = read_csv("./data_2/fivethirtyeight_datasets/snp.csv", show_col_types = FALSE) %>%
  janitor::clean_names() %>%
  separate(date,c("A", "B", "C")) %>%
  select("B","C","A", everything()) %>%
  rename(month = "B", day = "C", year = "A") %>%
  select(-day)%>%
  mutate(year = as.integer(year)) %>%
  mutate(month = month.abb[as.numeric(month)])%>%
  mutate(year = if_else(year < 23, year + 2000, year +1900)) %>%
  mutate(year = as.character(year)) %>%
  mutate(month = as.character(month)) %>%
  select(year,month,everything())
```

# Upload Unemployed dataset 
```{r}
unemp = read_csv("./data_2/fivethirtyeight_datasets/unemployment.csv", show_col_types = FALSE)
```

# Clean Unemployed dataset 
```{r}
unemp_2 = read_csv("./data_2/fivethirtyeight_datasets/unemployment.csv", show_col_types = FALSE) %>%
  janitor::clean_names()
```

# Change formate of Unemployed dataset

```{r}
unemp_3 = pivot_longer(
    unemp_2, 
    jan:dec,
    names_to = "month", 
    values_to = "percent_unemployment"
    ) %>%
  mutate(year = as.character(year))

```

# Left join SPN and Pols 

```{r}
merge_snp_pols = 
  left_join(snp_2,pols_month_2, by = c("year","month"))
```

# Left join merge to umployment 

```{r}
final_merge = 
  left_join(merge_snp_pols, unemp_3, by = c("year", "month")) 

```

```{r}
final_merge %>%
  pull(year) %>%
  range(na.rm = TRUE) 
```


The **pols month** dataset originally had `r nrow(pols_month)` observations and `r ncol(pols_month)` variables. Key variables before cleaning was **mon** and **prez_gop**, **prez_dem**,**sen_gop** and other variables that characterize political parties. **Mon** variable was a date that was written as "mon/day/year". Cleaned this variable by separating the one **mon** variable into three separate variables: **month**, **day** and **year**. Another way we cleaned this dataset was by deleting **date**, **prez_gop**, **prez_dem** and creating a new variable **president** that used infromation from prez_gop and prez_dem. 

The **snp** dataset originally had `r nrow(snp)` observations and `r ncol(snp)` variables.**Snp** only had two variables, **date** and close. **Date** variable took the form of "mon/day/year", however, ever cleaning this variable was separated into three separate variables, **month** , **day** and **year**. 

The **unemployment** dataset originally had `r nrow(unemp)` observations and `r ncol(unemp)` variables. The dataset was originally wide it needed to be cleaned so that it could be long format instead with three variables **year**, **month** and **percent unemployment**. 

The **final_merge** dateset after the proper joining, has `r nrow(final_merge)` observations and `r ncol(final_merge)` variables.The years range from `r final_merge %>% pull(year) %>% range(na.rm = TRUE)` years. Some of the key variables in **final merge** dataset are **year**, **month**, **percent unemployment** and some variables that characterize political parties. 


